---
title:
  '01 Microglia GO Ontology - `r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/"))) - 1]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile,
    encoding = encoding,
    output_file = file.path("..", "results", "R", "01-MG-GO.html")
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [ISS Microglia](https://github.com/eturkes/iss-microglia).*

The data here is derived from @`r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/"))) - 1]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/"))) - 1]``.

```{r}
#    This file is part of iss-microglia.
#    Copyright (C) 2020  Emir Turkes, Sebastiaan De Schepper, UK DRI at UCL
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# This section should be checked per document.
# --------------------------------------------
options(stringsAsFactors = FALSE)
packages <- c(
  "conflicted", "RColorBrewer", "GO.db", "GOSemSim", "ComplexHeatmap", "clusterProfiler", "cluster",
  "factoextra"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))

nbclust_fun <- function(x, k) {
  list(cluster = cutree(hclust(get_dist((x), "spearman"), "ward.D2"), k))
}
`%notin%` <- Negate(`%in%`)

analysis_no <- "01"
color <- colorRampPalette(rev(brewer.pal(9, "RdBu")))(100)
# --------------------------------------------

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------
data_name <- unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))] # Name of data.
assets_dir <- file.path("..", "assets") # Backed up objects and data.
cache_dir <- file.path("..", "cache", "R")
results_dir <- file.path("..", "results", "R")

if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 7)
# ----------------------------------------------------------
```

# Pilot

```{r}
run <- "pilot"

genes <- read.table(
  file.path(assets_dir, "gene-sets", paste0("MG-gene-symbols-", run, ".txt")), FALSE, "\t"
)
genes <- bitr(genes$V1, "SYMBOL", "ENTREZID", "org.Mm.eg.db")

rds <- file.path(cache_dir, paste0(analysis_no, "_", run, "_go_data.rds"))
if (file.exists(rds)) {
  go_data <- readRDS(rds)
} else {
  go_data <- godata("org.Mm.eg.db", "SYMBOL", "BP", TRUE)
  saveRDS(go_data, rds)
}

rds <- file.path(cache_dir, paste0(analysis_no, "_", run, "_sem_sim.rds"))
if (file.exists(rds)) {
  sem_sim <- readRDS(rds)
} else {
  sem_sim <- mgeneSim(genes$SYMBOL, go_data, "Jiang", NULL, verbose = FALSE)
  saveRDS(sem_sim, rds)
}
clustering <- hclust(get_dist(sem_sim, "spearman"), "ward.D2")

fviz_nbclust(
  sem_sim, nbclust_fun, k.max = nrow(sem_sim) / 2, print.summary = FALSE, verbose = FALSE
)
fviz_nbclust(
  sem_sim, nbclust_fun, "wss", k.max = nrow(sem_sim) / 2, print.summary = FALSE, verbose = FALSE
)

rds <- file.path(cache_dir, paste0(analysis_no, "_", run, "_nbclust.rds"))
if (file.exists(rds)) {
  nbclust <- readRDS(rds)
} else {
  nbclust <- clusGap(
    sem_sim, nbclust_fun, nrow(sem_sim) / 2, d.power = 2, verbose = FALSE
  )
  saveRDS(nbclust, rds)
}
split <- maxSE(nbclust$Tab[ , "gap"], nbclust$Tab[ , "SE.sim"], "globalSEmax")
fviz_gap_stat(nbclust, maxSE = list(method = "globalSEmax"))

set.seed(1)
heatmap <- Heatmap(
  sem_sim, color, row_split = split, column_split = split,
  show_heatmap_legend = FALSE, cluster_rows = clustering, cluster_columns = clustering
)
draw(heatmap)

clusters <- row_order(heatmap)
genes_sub <- genes[genes$SYMBOL %in% rownames(sem_sim), ]
for (i in seq_along(clusters)) {
  if (length(clusters[[i]]) > 1) {
    clusters[[i]] <- genes_sub$ENTREZID[row_order(heatmap)[[i]]]
    names(clusters)[i] <- paste0("c", i)
  }
}
if (any(is.na(names(clusters)))) {
  clusters <- clusters[-which(is.na(names(clusters)))]
}

sem_sub <- sem_sim[unlist(row_order(heatmap)), unlist(column_order(heatmap))]
genes_sub <- genes[genes$ENTREZID %in% unlist(clusters), ]
sem_sub <- sem_sim[rownames(sem_sim) %in% genes_sub$SYMBOL, colnames(sem_sim) %in% genes_sub$SYMBOL]
not_clustered <- genes$SYMBOL[genes$SYMBOL %notin% rownames(sem_sub)]
print(
  paste0(
    "Genes not clustered (", length(not_clustered), "): ", paste(not_clustered, collapse = ", ")
  )
)

rds <- file.path(cache_dir, paste0(analysis_no, "_", run, "_compare_clusters.rds"))
if (file.exists(rds)) {
  compare_clusters <- readRDS(rds)
} else {
  compare_clusters <- compareCluster(clusters, OrgDb = "org.Mm.eg.db", ont = "BP")
  saveRDS(compare_clusters, rds)
}

dotplot(compare_clusters)
```

# NLF

```{r}
run <- "NLF"

genes <- read.table(
  file.path(assets_dir, "gene-sets", paste0("MG-gene-symbols-", run, ".txt")), FALSE, "\t"
)
genes <- bitr(genes$V1, "SYMBOL", "ENTREZID", "org.Mm.eg.db")

rds <- file.path(cache_dir, paste0(analysis_no, "_", run, "_go_data.rds"))
if (file.exists(rds)) {
  go_data <- readRDS(rds)
} else {
  go_data <- godata("org.Mm.eg.db", "SYMBOL", "BP", TRUE)
  saveRDS(go_data, rds)
}

rds <- file.path(cache_dir, paste0(analysis_no, "_", run, "_sem_sim.rds"))
if (file.exists(rds)) {
  sem_sim <- readRDS(rds)
} else {
  sem_sim <- mgeneSim(genes$SYMBOL, go_data, "Jiang", NULL, verbose = FALSE)
  saveRDS(sem_sim, rds)
}
clustering <- hclust(get_dist(sem_sim, "spearman"), "ward.D2")
```

```{r, fig.width = 25, fig.height = 25}
fviz_nbclust(
  sem_sim, nbclust_fun, k.max = nrow(sem_sim) / 2, print.summary = FALSE, verbose = FALSE
)
fviz_nbclust(
  sem_sim, nbclust_fun, "wss", k.max = nrow(sem_sim) / 2, print.summary = FALSE, verbose = FALSE
)

rds <- file.path(cache_dir, paste0(analysis_no, "_", run, "_nbclust.rds"))
if (file.exists(rds)) {
  nbclust <- readRDS(rds)
} else {
  nbclust <- clusGap(
    sem_sim, nbclust_fun, nrow(sem_sim) / 2, d.power = 2, verbose = FALSE
  )
  saveRDS(nbclust, rds)
}
split <- maxSE(nbclust$Tab[ , "gap"], nbclust$Tab[ , "SE.sim"], "globalSEmax", 2)
fviz_gap_stat(nbclust, maxSE = list(method = "globalSEmax", SE.factor = 2))

split <- 9
set.seed(1)
heatmap <- Heatmap(
  sem_sim, color, column_names_gp = gpar(fontsize = 10),
  row_names_gp = gpar(fontsize = 10), cluster_rows = clustering, cluster_columns = clustering,
  row_split = split, column_split = split, show_heatmap_legend = FALSE,
  row_title_gp = gpar(fontsize = 12), column_title_gp = gpar(fontsize = 12)
)
draw(heatmap)

clusters <- row_order(heatmap)
genes_sub <- genes[genes$SYMBOL %in% rownames(sem_sim), ]
for (i in seq_along(clusters)) {
  if (length(clusters[[i]]) > 1) {
    clusters[[i]] <- genes_sub$ENTREZID[row_order(heatmap)[[i]]]
    names(clusters)[i] <- paste0("c", i)
  }
}
if (any(is.na(names(clusters)))) {
  clusters <- clusters[-which(is.na(names(clusters)))]
}

sem_sub <- sem_sim[unlist(row_order(heatmap)), unlist(column_order(heatmap))]
genes_sub <- genes[genes$ENTREZID %in% unlist(clusters), ]
sem_sub <- sem_sim[rownames(sem_sim) %in% genes_sub$SYMBOL, colnames(sem_sim) %in% genes_sub$SYMBOL]
not_clustered <- genes$SYMBOL[genes$SYMBOL %notin% rownames(sem_sub)]
print(
  paste0(
    "Genes not clustered (", length(not_clustered), "): ", paste(not_clustered, collapse = ", ")
  )
)
```

```{r, fig.width = 15, fig.height = 12}
rds <- file.path(cache_dir, paste0(analysis_no, "_", run, "_compare_clusters.rds"))
if (file.exists(rds)) {
  compare_clusters <- readRDS(rds)
} else {
  compare_clusters <- compareCluster(clusters, OrgDb = "org.Mm.eg.db", ont = "BP")
  saveRDS(compare_clusters, rds)
}

dotplot(compare_clusters)
```

# References

This is the concluding section of the document.
Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
