---
title:
  '01 Microglia GO Ontology - `r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/"))) - 1]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile,
    encoding = encoding,
    output_file = file.path("..", "results", "R", "01-MG-GO.html")
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [ISS Microglia](https://github.com/eturkes/iss-microglia).*

The data here is derived from @`r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/"))) - 1]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/"))) - 1]``.

```{r}
#    This file is part of iss-microglia.
#    Copyright (C) 2020  Emir Turkes
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# This section should be checked per document.
# --------------------------------------------
packages <- c("conflicted", "scater", "plotly", "scran", "pheatmap", "viridis")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))

analysis_no <- 1
# --------------------------------------------

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------
data_name <- unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))] # Name of data.
assets_dir <- file.path("..", "assets") # Backed up objects and data.
cache_dir <- file.path("..", "cache", "R")
results_dir <- file.path("..", "results", "R")

if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 7)
# ----------------------------------------------------------
```

# Analysis

```{r}
quickgo <- read.table(file.path(assets_dir, "gene-sets", "quickgo-BP.tsv"), TRUE, "\t")
quickgo <- quickgo[ , c(6, 3)]
mat <- as.data.frame.matrix(table(quickgo))
mat <- as.matrix(mat)
mat[mat > 0] = 1
genes <- data.frame(genes = colnames(mat))

sce <- SingleCellExperiment(list(logcounts = mat), colData = genes)
set.seed(1)
sce <- runPCA(sce, ncomponents = 2)
ggplotly(plotPCA(sce, colour_by = "genes"))

set.seed(1)
g <- buildSNNGraph(sce, 2, use.dimred = "PCA")
clusters <- igraph::cluster_walktrap(g)$membership
sce$clusters <- as.factor(clusters)
paste0(paste0("cluster", sce$clusters), "-", sce$genes)[order(
  paste0(paste0("cluster", sce$clusters), "-", sce$genes)
)]
ggplotly(plotPCA(sce, colour_by = "clusters"))

markers <- findMarkers(
  sce, groups = sce$clusters, direction = "up", test.type = "wilcox"
)

markers_sub <- markers[[1]]
markers_sub <- markers_sub[markers_sub$Top <= 5, ]
AUC <- markers_sub[ , c(4:ncol(markers_sub))]
pheatmap(
  AUC, viridis(21), breaks = seq(0, 1, length.out = 21), cluster_cols = FALSE, main = "Cluster 1"
)

markers_sub <- markers[[2]]
markers_sub <- markers_sub[markers_sub$Top <= 5, ]
AUC <- markers_sub[ , c(4:ncol(markers_sub))]
pheatmap(
  AUC, viridis(21), breaks = seq(0, 1, length.out = 21), cluster_cols = FALSE, main = "Cluster 2"
)
markers_sub <- markers[[3]]
markers_sub <- markers_sub[markers_sub$Top <= 5, ]
AUC <- markers_sub[ , c(4:ncol(markers_sub))]
pheatmap(
  AUC, viridis(21), breaks = seq(0, 1, length.out = 21), cluster_cols = FALSE, main = "Cluster 3"
)
markers_sub <- markers[[4]]
markers_sub <- markers_sub[markers_sub$Top <= 5, ]
AUC <- markers_sub[ , c(4:ncol(markers_sub))]
pheatmap(
  AUC, viridis(21), breaks = seq(0, 1, length.out = 21), cluster_cols = FALSE, main = "Cluster 4"
)
markers_sub <- markers[[5]]
markers_sub <- markers_sub[markers_sub$Top <= 5, ]
AUC <- markers_sub[ , c(4:ncol(markers_sub))]
pheatmap(
  AUC, viridis(21), breaks = seq(0, 1, length.out = 21), cluster_cols = FALSE, main = "Cluster 5"
)
```

# References

This is the concluding section of the document.
Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
