---
title:
  '02 Cluster Proportions - `r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/"))) - 1]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile,
    encoding = encoding,
    output_file = file.path("..", "results", "R", "02-cluster-proportion.html")
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [ISS Microglia](https://github.com/eturkes/iss-microglia).*

The data here is derived from @`r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/"))) - 1]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/"))) - 1]``.

```{r}
#    This file is part of iss-microglia.
#    Copyright (C) 2020  Emir Turkes, Sebastiaan De Schepper, UK DRI at UCL
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# This section should be checked per document.
# --------------------------------------------
options(stringsAsFactors = FALSE)
packages <- c("conflicted", "ggplot2")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source("utils.R")

analysis_no <- "02"
# --------------------------------------------

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------
data_name <- unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))] # Name of data.
assets_dir <- file.path("..", "assets") # Backed up objects and data.
cache_dir <- file.path("..", "cache", "R")
results_dir <- file.path("..", "results", "R")

if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 7)
# ----------------------------------------------------------
```

# Prep

```{r}
gene_names <- read.table(
  file.path(assets_dir, "gene-sets", paste0("MG-gene-symbols-NLF", ".txt")), FALSE, "\t"
)
gene_names$V2 <- rownames(gene_names)

cluster_list <- vector("list", 6)
for (i in seq_along(cluster_list)) {
  cluster_list[[i]] <- read.table(
    file.path(results_dir, "new-clusters", paste0("group", i, ".txt")), FALSE, "\t"
  )
}

gene_names$V3 <- gene_names$V2
for (i in seq_along(gene_names$V3)) {
  gene_names$V3[i] <- which(
    sapply(seq_along(cluster_list), function(x) {gene_names$V1[i] %in% cluster_list[[x]]$V1})
  )
}
```

# CTRL

```{r, fig.height = 20}
sample <- "CTRL"

spot_codes <- read.delim(
  file.path(assets_dir, "codebooks", "NLF-150genes", paste0(sample, "-SpotCode.txt")), FALSE
)
spot_names <- merge(spot_codes, gene_names, by.x = "V1", by.y = "V2")
colnames(spot_names) <- c("spot_code", "gene_names", "cluster")

spot_names_table <- as.data.frame(table(spot_names$gene_names))
spot_names_table <- merge(spot_names_table, spot_names, by.x = "Var1", by.y = "gene_names")
spot_names_table <- spot_names_table[!duplicated(spot_names_table), ]
colnames(spot_names_table) <- c("gene_names", "freq", "spot_code", "cluster")
spot_names_table$gene_names <- factor(spot_names_table$gene_names, rev(unlist(cluster_list)))

ggplot(spot_names_table, aes(x = gene_names, y = freq, fill = cluster)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    x = "Genes", y = "Number of Spots (Log Scale)",
    title = paste0(sample, " Genes vs. Number of Spots")
  ) +
  theme_black() +
  scale_y_log10(n.breaks = 6, expand = expansion(mult = c(0, 0)), limits = c(1, 20000)) +
  theme(
    panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank(),
    legend.key.height = unit(3, "cm"), axis.title.x = element_text(size = 25),
    axis.title.y = element_text(size = 25), axis.text.x = element_text(size = 20),
    plot.title = element_text(size = 30), legend.title = element_text(size = 25),
    legend.text = element_text(size = 20)
  ) +
  scale_fill_manual(
    name = "Clusters",
    labels = c(
      "Interferon signaling\nAntigen processing (MHC class II)\nMyeloid leukocyte activation",
      "ATPase activity\npH regulation",
      "Lysosome activity\nProtein processing\nPeptidase activity",
      "Phagocytic cup formation\nFc-gamma receptor activity\nActin cytoskeletal processes",
      "Phagosome formation\nActin cytoskeletal processes",
      "Uncatagorized"
    ),
    values = c(
      hsv(1, 1, 1), hsv(0.13, 1, 1), hsv(1 / 3, 1, 1),
      hsv(0.1, 1, 0.6), hsv(0.55, 1, 1), hsv(0, 0, 1)
    )
  )
```

# CTRL

```{r, fig.height = 20}
sample <- "NLF"

spot_codes <- read.delim(
  file.path(assets_dir, "codebooks", "NLF-150genes", paste0(sample, "-SpotCode.txt")), FALSE
)
spot_names <- merge(spot_codes, gene_names, by.x = "V1", by.y = "V2")
colnames(spot_names) <- c("spot_code", "gene_names", "cluster")

spot_names_table <- as.data.frame(table(spot_names$gene_names))
spot_names_table <- merge(spot_names_table, spot_names, by.x = "Var1", by.y = "gene_names")
spot_names_table <- spot_names_table[!duplicated(spot_names_table), ]
colnames(spot_names_table) <- c("gene_names", "freq", "spot_code", "cluster")
spot_names_table$gene_names <- factor(spot_names_table$gene_names, rev(unlist(cluster_list)))

ggplot(spot_names_table, aes(x = gene_names, y = freq, fill = cluster)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    x = "Genes", y = "Number of Spots (Log Scale)",
    title = paste0(sample, " Genes vs. Number of Spots")
  ) +
  theme_black() +
  scale_y_log10(n.breaks = 6, expand = expansion(mult = c(0, 0)), limits = c(1, 20000)) +
  theme(
    panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank(),
    legend.key.height = unit(3, "cm"), axis.title.x = element_text(size = 25),
    axis.title.y = element_text(size = 25), axis.text.x = element_text(size = 20),
    plot.title = element_text(size = 30), legend.title = element_text(size = 25),
    legend.text = element_text(size = 20)
  ) +
  scale_fill_manual(
    name = "Clusters",
    labels = c(
      "Interferon signaling\nAntigen processing (MHC class II)\nMyeloid leukocyte activation",
      "ATPase activity\npH regulation",
      "Lysosome activity\nProtein processing\nPeptidase activity",
      "Phagocytic cup formation\nFc-gamma receptor activity\nActin cytoskeletal processes",
      "Phagosome formation\nActin cytoskeletal processes",
      "Uncatagorized"
    ),
    values = c(
      hsv(1, 1, 1), hsv(0.13, 1, 1), hsv(1 / 3, 1, 1),
      hsv(0.1, 1, 0.6), hsv(0.55, 1, 1), hsv(0, 0, 1)
    )
  )
```

# References

This is the concluding section of the document.
Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
